---
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    css: "brand_styles.css"
    includes:
      before_body: "before_body.html"
    self_contained: false
---

```{r intro-block, results='asis', echo=FALSE}

htmltools::div(
  style = "
    transform: scale(0.9);
    transform-origin: top left;
    width: 111%;
  ",
  
  htmltools::div(
    style = "
      position:relative;
      margin-top:18px;
      display:grid;
      grid-template-columns: 2.2fr 1fr;
      gap:20px;
      align-items:stretch;
    ",
    
  # ================= LEFT PANEL =================
  htmltools::div(
style = "
  background: rgba(25, 95, 170, 0.55);
  backdrop-filter: blur(14px);
  -webkit-backdrop-filter: blur(14px);
  border: 1px solid rgba(255,255,255,0.16);
  border-radius:18px;
  overflow:hidden;
  box-shadow:0 6px 20px rgba(0,0,0,0.18);
  color:white;
  display:flex;
  flex-direction:column;
",

    # ---------- BANNER IMAGE ----------
    htmltools::tags$img(
      src = "FB TRPP.png",
      style = "
        width:100%;
        height:220px;
        object-fit:cover;
        display:block;
      "
    ),

    # ---------- TEXT CONTENT ----------
    htmltools::div(
      style = "
        padding:22px 24px 24px 24px;
        font-size:15px;
        line-height:1.65;
        color:rgba(255,255,255,0.95);
      ",

      htmltools::tags$p(
        htmltools::a(
          href = "https://www.therockpoolproject.co.uk/the-big-rock-pool-challenge",
          style = "
            color:#D6F2FF;
            font-weight:700;
            text-decoration:none;
          ",
          "The Big Rock Pool Challenge"
        ),
        " is a UK-wide citizen science initiative run by the Rock Pool Project, bringing people together to explore and record life on UK rocky shores. Participants take part in monthly BioBlitz Battle games, recording coastal marine life at designated monitoring hubs using the iNaturalist platform and the BioBlitz Battle app."
      ),

      htmltools::tags$p(
        "Records are verified through community identifications, allowing observations to reach research-grade status. Every record added helps build a growing picture of coastal biodiversity around the UK."
      ),

      htmltools::tags$p(
        "The Big Rock Pool Challenge is run by The Rock Pool Project, in partnership with the Marine Biological Association, and is funded by the ScottishPower Foundation and the National Lottery Heritage Fund."
      )
    )
  ),

  # ================= RIGHT PANEL =================
  htmltools::div(
    style = "
      background:#ffffff;
      color:#222222;
      border-radius:18px;
      padding:20px;
      display:flex;
      flex-direction:column;
      justify-content:center;
      align-items:center;
      gap:24px;
      box-shadow:0 6px 18px rgba(0,0,0,0.10);
    ",

    htmltools::tags$div(
      "Funding & Partners",
      style = "
        font-size:15px;
        font-weight:700;
        letter-spacing:0.3px;
        color:#222222;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/67dc251080bf058c3bffd547_hf-logo.png",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/690234a2aeabe1446d63c1cc_AAFF%20LOGOTYPE%20SCOTTISHPOWER%20FOUNDATION%20RGB-p-2000.jpg",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/67b4f7fae1b32ae21a76b697_Untitled%20design%20(11).png",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    )
  )
)
)

```

::: {style="height:20px;"}
:::
```{=html}
<script>
window.addEventListener("load", function() {

  function moveTOC() {

    var toc = document.getElementById("TOC");
    var target = document.getElementById("toc-here");

    if (toc && target && !target.contains(toc)) {
      target.appendChild(toc);
      toc.style.marginTop = "20px";
    } else {
      setTimeout(moveTOC, 100);
    }
  }

  moveTOC();
});
</script>
```

<div id="toc-here"></div>

```{r setup-chunk, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(sf)
library(tidyr)
library(vegan)
library(ggrepel)
library(scales)
library(forcats)
library(glue)
library(purrr)
library(rnaturalearth)
library(viridis)
library(htmltools)
library(ggimage)

theme_set(
  theme_minimal() +
    theme(
      legend.position = "right",
      plot.caption = element_text(hjust = 0)
    )
)

safe_id <- function(x) {
  gsub("[^A-Za-z0-9]", "-", x)
}

get_col <- function(df, nm) {
  if (nm %in% names(df)) df[[nm]] else rep(NA_character_, nrow(df))
}

default_avatar <- "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/68c02913eb9eedda1582018a_iNaturalist%20app.png"

safe_img <- function(x) {
  ifelse(is.na(x) | x == "", default_avatar, x)
}

safe_species_image <- function(photo_url,
                               attribution,
                               license_ok = TRUE) {

  ok <- !is.na(photo_url) &
        photo_url != "" &
        license_ok

  out_photo <- ifelse(ok, photo_url, default_avatar)

  out_attr <- ifelse(ok, attribution, "")

  tibble(
    photo = out_photo,
    attribution = out_attr
  )
}

allowed_licenses <- c(
  "cc-by",
  "cc-by-sa",
  "cc-by-nc",
  "cc-by-nc-sa",
  "cc0"
)


```

```{r data-download, echo=FALSE}

source("../../Ben/iNat API functions.R")

iconic_lookup <- read.csv("../../Data/BRPC_iconic_taxa.csv")

objs <- load("../../Data/rarity_scores2.1.RData")
rarity_scores <- get(objs[1])

rarity_scores <- rarity_scores %>%
  select(-Score) %>%
  rename(Score = Score100)

load("../../Ben/BRPC_dat.RData")
```

```{r hub-setup}

hub_sf <- tibble::tribble(
  ~hub, ~lon, ~lat,
  "Castle Beach", -5.06, 50.15,
  "Lee-on-the-Solent", -1.20, 50.80,
  "Beadnell Haven", -1.63, 55.56,
  "Ovingdean", -0.08, 50.81,
  "Shoeburyness", 0.78, 51.53,
  "Menai Bridge", -4.16, 53.23,
  "Mount Batten", -4.14, 50.36,
  "Kilve Beach", -3.15, 51.20,
  "East Sands", -2.79, 56.34
) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)
```

```{r data-pipeline}

# ---------------------------------------
# ALL DATA
# ---------------------------------------

BRPC_dat <- BRPC_dat %>%
  mutate(
    scientific_name = dplyr::coalesce(
      get_col(., "scientific_name"),
      get_col(., "taxon.name"),
      get_col(., "taxon_name"),
      get_col(., "species_guess")
    )
  )
# ---------------------------------------
# PROJECT DATA (up to Q1 end)
# ---------------------------------------

project_end <- as.Date("2026-03-31")

BRPC_dat <- BRPC_dat %>%
  mutate(observed_on = as.Date(observed_on)) %>%
  filter(observed_on <= project_end)

all_sf <- BRPC_dat %>%
  mutate(
    lon = purrr::map_dbl(
      geojson.coordinates,
      ~ if(length(.x) >= 1) .x[[1]] else NA_real_
    ),
    lat = purrr::map_dbl(
      geojson.coordinates,
      ~ if(length(.x) >= 2) .x[[2]] else NA_real_
    )
  ) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)

all_sf$hub <- hub_sf$hub[
  st_nearest_feature(all_sf, hub_sf)
]

BRPC_dat <- all_sf %>%
  st_drop_geometry()

# ---------------------------------------
# Q1 DATA
# ---------------------------------------

q1_base <- BRPC_dat %>%
  filter(
    observed_on >= "2026-01-01",
    observed_on <= "2026-03-31"
  )

q1_sf <- q1_base %>%
  mutate(
    lon = purrr::map_dbl(geojson.coordinates, 1),
    lat = purrr::map_dbl(geojson.coordinates, 2)
  ) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)

q1_sf$hub <- hub_sf$hub[
  st_nearest_feature(q1_sf, hub_sf)
]

q1_base$hub <- q1_sf$hub
```

```{r helper-functions}

render_summary_card <- function(title, observations, contributors, species) {

  div(
    style = "
      background:#ffffff;
      border-radius:20px;
      padding:26px;
      margin:20px 0 30px 0;
      box-shadow:0 4px 14px rgba(0,0,0,0.08);
      border:1px solid #e8e8e8;
      text-align:center;
    ",

    h2(title),

    div(
      style = "
        display:inline-flex;
        gap:18px;
        flex-wrap:wrap;
        justify-content:center;
        align-items:center;
        background:#f5f5f5;
        padding:14px 22px;
        border-radius:999px;
        margin-top:10px;
      ",

      HTML(glue("
        <span><strong>{comma(observations)}</strong> observations</span>
        <span>·</span>
        <span><strong>{comma(contributors)}</strong> contributors</span>
        <span>·</span>
        <span><strong>{comma(species)}</strong> species</span>
      "))
    )
  )
}

icon_lookup <- tibble::tribble(
  ~iconic_group,        ~icon_path,
  "Brown algae",        "icons/Brown algae.png",
  "Crustaceans",        "icons/Crustaceans.png",
  "Ray-finned Fishes",  "icons/Fishes.png",
  "Green algae",        "icons/Green algae.png",
  "Molluscs",           "icons/Molluscs.png",
  "Red algae",          "icons/Red algae.png",
  "Segmented Worms",    "icons/Segmented worms.png",
  "Sponges",            "icons/Sponges.png",
  "Echinoderms",        "icons/Starfish.png",
  "Tunicates",          "icons/Tunicates.png",
  "Sea Anemones",           "icons/Anemones.png"
)
```

# Big Rock Pool Challenge Biodiversity Report - Q1 2026

This report brings together all observations recorded through the Big Rock Pool Challenge up to March 2026, with a closer look at activity during Quarter 1 of 2026 (January–March).

## All project data

The Big Rock Pool Challenge commenced in October 2024 at hubs located at Castle Beach, Falmouth and Mount Batten, Plymouth. In October 2025, seven new hubs joined the Big Rock Pool Challenge:

- Shoeburyness, Southend-on-Sea
- Kilve Beach, Somerset
- Beadnell Haven, Northumberland
- East Sands, St Andrews
- Lee-on-the-Solent, Hampshire
- Menai Bridge, Bangor
- Ovingdean, Brighton

Data presented in this section of the report reflect all Big Rock Pool Challenge records from the start of the project in October 2024, until the end of Quarter 1 (end of March) 2026.

**To view data from Quarter 1 of 2026 only, please see the second half of this report, or [click here](#q1-2026).**

```{r overall-summary}

tagList(

  render_summary_card(
    "Overall Project Summary",
    observations = nrow(BRPC_dat),
    contributors = n_distinct(BRPC_dat$user.login),
    species = n_distinct(na.omit(BRPC_dat$scientific_name))
  )
)

```

Species are grouped into higher-level “iconic taxa” categories based on their taxonomic ancestry. This grouping simplifies complex species-level data into broader biological categories (e.g. molluscs, crustaceans, algae), allowing clearer interpretation of community structure across sites and time periods.

```{r iconic-taxa}

all_iconic <- BRPC_dat %>%
  mutate(
    ancestry_vec = strsplit(
      ifelse(is.na(taxon.ancestry), "", taxon.ancestry),
      "/"
    ),

    iconic_group = purrr::map_chr(ancestry_vec, function(x) {

      x <- suppressWarnings(as.integer(x))

      hit <- iconic_lookup$Common[
        iconic_lookup$iNat_code %in% x
      ]

      if (length(hit) == 0) "Other" else hit[1]
    })
  )

library(ggtext)

all_iconic_summary <- all_iconic %>%
  count(iconic_group, sort = TRUE)

all_iconic_summary <- all_iconic_summary %>%
  left_join(icon_lookup, by = "iconic_group")

ggplot(
  all_iconic_summary,
  aes(
    x = reorder(iconic_group, n),
    y = n
  )
) +

  geom_col(
    fill = "#9ecae1",
    width = 0.72
  ) +

  coord_flip() +

  scale_y_continuous(labels = comma) +

  geom_image(
    aes(image = icon_path),
    by = "width",
    size = 0.06,
    nudge_y = max(all_iconic_summary$n) * 0.03
  ) +

  expand_limits(
    y = max(all_iconic_summary$n) * 1.12
  ) +

  labs(
    title = "Iconic Taxa Breakdown Across All Project Data",
    x = NULL,
    y = "Observations"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 11),
    plot.title = element_text(face = "bold")
  )

```

```{r q1-iconic-taxa}

q1_iconic <- q1_base %>%
  mutate(
    ancestry_vec = strsplit(
      ifelse(is.na(taxon.ancestry), "", taxon.ancestry),
      "/"
    ),

    iconic_group = purrr::map_chr(ancestry_vec, function(x) {

      x <- suppressWarnings(as.integer(x))

      hit <- iconic_lookup$Common[
        iconic_lookup$iNat_code %in% x
      ]

      if (length(hit) == 0) "Other" else hit[1]
    })
  )

iconic_summary <- q1_iconic %>%
  count(iconic_group, sort = TRUE)

iconic_summary <- iconic_summary %>%
  left_join(icon_lookup, by = "iconic_group")


```

```{r add-rarity-scores}

# ---------------------------------------
# Join rarity scores
# ---------------------------------------

all_iconic <- all_iconic %>%
  left_join(
    rarity_scores,
    by = c("scientific_name" = "name")
  )

q1_iconic <- q1_iconic %>%
  left_join(
    rarity_scores,
    by = c("scientific_name" = "name")
  )
```

The Big Rock Pool Challenge is powered by people. Every observation and identification is a contribution to a shared understanding of coastal biodiversity.

Some participants frequently attend events and submit records, while others help out by identifying species through verification on iNaturalist- all contributions are equally valuable in building the dataset.

```{r overall-contributor-summary, results='asis', echo=FALSE}

recorders_n <- n_distinct(BRPC_dat$user.login)

verifiers_n <- BRPC_dat %>%
  filter(!is.na(identifications)) %>%
  pull(identifications) %>%
  purrr::map_dfr(~{
    if (is.null(.x) || !"user.login" %in% names(.x)) return(NULL)
    tibble(user = .x$user.login)
  }) %>%
  filter(!is.na(user), user != "") %>%
  distinct(user) %>%
  nrow()

HTML(glue::glue(
  "<p style='text-align:left;font-size:16px;margin:10px 0 20px 0;'>
  For all project data until end of March 2026, records were submitted by <strong>{recorders_n}</strong> recorders,
  and verified to research grade by <strong>{verifiers_n}</strong> identifiers.
  </p>"
))
```

The top recorders and identifiers are presented below- thank you for your efforts!

```{r overall-top-recorders-panels, results='asis', echo=FALSE}
library(dplyr)
library(htmltools)
library(scales)

default_avatar <- "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/68c02913eb9eedda1582018a_iNaturalist%20app.png"

overall_recorders <- BRPC_dat %>%
  filter(!is.na(user.login)) %>%
  group_by(user.login) %>%
  summarise(
    records = n(),
    icon = first(na.omit(user.icon_url)),
    .groups = "drop"
  ) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    records = comma(records),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",
paste0(
  "<a href='", icon, "' target='_blank'>",
  "<div style='
    width:64px;
    height:64px;
    border-radius:50%;
    margin:0 auto 10px auto;
    background-image:url(\"", icon, "\");
    background-size:cover;
    background-position:center;
    border:2px solid #ececec;
  '></div>",
  "</a>"
),

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      records,
      " records</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Overall Top Recorders", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(overall_recorders$card_html, collapse = ""))
  )
)
```

```{r overall-top-identifiers-panels, results='asis', echo=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(htmltools)
library(scales)

overall_identifiers <- BRPC_dat %>%
  filter(!is.na(identifications)) %>%
  select(identifications) %>%
  mutate(
    identifier = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.login" %in% names(.x)) {
        character(0)
      } else {
        .x$user.login
      }
    }),
    icon = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.icon_url" %in% names(.x)) {
        character(0)
      } else {
        .x$user.icon_url
      }
    })
  ) %>%
  select(identifier, icon) %>%
  unnest(c(identifier, icon)) %>%
  filter(!is.na(identifier), identifier != "") %>%
  group_by(identifier) %>%
  summarise(
    identifications = n(),
    icon = first(na.omit(icon)),
    .groups = "drop"
  ) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    identifications = comma(identifications),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",

     paste0(
  "<a href='", icon, "' target='_blank'>",
  "<div style='
    width:64px;
    height:64px;
    border-radius:50%;
    margin:0 auto 10px auto;
    background-image:url(\"", icon, "\");
    background-size:cover;
    background-position:center;
    border:2px solid #ececec;
  '></div>",
  "</a>"
),

      "<div style='font-weight:600;font-size:14px;'>",
      identifier,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      identifications,
      " identifications</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Overall Top Identifiers", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(overall_identifiers$card_html, collapse = ""))
  )
)
```

::: {style="height:20px;"}
:::

Each hub represents a local coastal site where volunteers return to record marine life over time. These locations form the backbone of the Big Rock Pool Challenge network.

Because each hub has been running for different lengths of time, and is explored by different groups of people, the number of records varies, but every contribution adds valuable information about rock pool communities at that site.

Hubs are arranged in alphabetical order.

```{r map}

uk <- ne_countries(
  scale = "medium",
  country = "United Kingdom",
  returnclass = "sf"
) |>
  st_transform(3857)

hub_sf <- tibble::tribble(
  ~hub, ~lon, ~lat,
  "Castle Beach", -5.06, 50.15,
  "Lee-on-the-Solent", -1.20, 50.80,
  "Beadnell Haven", -1.63, 55.56,
  "Ovingdean", -0.08, 50.81,
  "Shoeburyness", 0.78, 51.53,
  "Menai Bridge", -4.16, 53.23,
  "Mount Batten", -4.14, 50.36,
  "Kilve Beach", -3.15, 51.20,
  "East Sands", -2.79, 56.34
) |>
  mutate(hub_id = row_number()) |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |>
  st_transform(3857)

q1_sf$hub <- hub_sf$hub[
  sf::st_nearest_feature(q1_sf, hub_sf)
]

q1_base$hub <- q1_sf$hub

all_data <- sf::st_drop_geometry(all_sf)

hub_summary <- all_data %>%
  group_by(hub) %>%
  summarise(
    records = n(),
    participants = n_distinct(user.login),
    species = n_distinct(scientific_name),
    .groups = "drop"
  )

hub_sf <- hub_sf %>%
  left_join(hub_summary, by = "hub")

ggplot() +

  geom_sf(
    data = uk,
    fill = "grey95",
    colour = "grey50"
  ) +

  geom_sf(
    data = q1_sf,
    colour = "grey70",
    alpha = 0.25,
    size = 0.3
  ) +

  geom_sf(
    data = hub_sf,
    aes(size = records),
    colour = "#1f78b4",
    fill = "#56B4E9",
    shape = 21,
    stroke = 1.2
  ) +

  geom_sf_text(
    data = hub_sf,
    aes(label = hub_id),
    colour = "black",
    fontface = "bold",
    size = 3
  ) +

 scale_size_area(
  max_size = 18,
  name = "Observations"
)+

  labs(
    title = "Big Rock Pool Challenge Hubs",
    subtitle = "Records as of 31/03/2026"
  ) +

  theme_void() +

  theme(
    legend.position = "right",

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(
      size = 11,
      margin = margin(b = 10)
    )
  )
```

```{r records-by-hub}

all_data_hub_summary <- BRPC_dat %>%
  group_by(hub) %>%
  summarise(
    observations = n(),
    contributors = n_distinct(user.login),
    species = n_distinct(scientific_name),
    .groups = "drop"
  ) %>%
  left_join(
    st_drop_geometry(hub_sf) %>%
      select(hub_id, hub),
    by = "hub"
  ) %>%
  select(
    Hub_ID = hub_id,
    Hub = hub,
    Observations = observations,
    Contributors = contributors,
    Species = species
  ) %>%
  arrange(Hub)

kable(all_data_hub_summary)
```

```{r hub-summary}
all_hub_summaries <- all_iconic %>%
  group_by(hub) %>%
  summarise(
    records = n(),
    contributors = n_distinct(user.login),
    species = n_distinct(scientific_name),

    top_observer = names(sort(table(user.login), decreasing = TRUE))[1],
    top_observer_n = max(table(user.login)),

    .groups = "drop"
  )
```

```{r hub-identifier-summary}

all_hub_top_identifiers <- all_iconic %>%
  filter(!is.na(identifications)) %>%
  select(hub, identifications) %>%
  tidyr::unnest(identifications) %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    identifications = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(

   icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(identifications),
      " identifications</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_identifiers_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )
```

```{r top-recorders}
all_hub_top_recorders <- all_iconic %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    records = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(

   icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(records),
      " observations</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_recorders_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )
```

```{r load-hub-logos}
hub_logos <- tibble::tribble(
  ~hub, ~logo_url,
  "Castle Beach", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e52dddfca602b7e70fd34f_Castle%20Beach_background.png",
  "Lee-on-the-Solent", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f7a42bf2ced5a97ff1d7cc_Lee-on-the-Solent_background.png",
  "Beadnell Haven", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68ee485ce2c8ba073e90f031_Beadnell%20Haven_background.png",
  "Ovingdean", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b2e3ee0622ac5288d0ea_Ovingdean%20Beach_background.png",
  "Shoeburyness", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/69023480819461010c5b7777_Shoeburyness_background.png",
  "Menai Bridge", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f66091a36dd2e5a7f195a6_Menai%20Bridge_background.png",
  "Mount Batten", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68ee4a28c6ce42f665792558_Mount%20Batten_background.png",
  "Kilve Beach", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7af1e8dbb099b8d5242b3_Kilve%20Beach_background.png",
  "East Sands", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b07fedeb13e53466355c_East%20Sands%20Beach_background.png"
)
```

```{r load-hub-images}
hub_site_photos <- tibble::tribble(
  ~hub, ~site_photo,

  "Mount Batten",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/687a3d7146a339fab83b27e5_Plymouth_lead_volunteers.jpg",

  "Castle Beach",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e525f5687223457755b1ef_4%20-%20Alexander%20Lydon.png",

  "Menai Bridge",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f65e11c5ec88584d09a9f8_menai%20bridge%20site.jpg",

  "Ovingdean",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b2c1af58aa97a264790e_IMG_3893%20-%20Sophie%20Crocker.jpeg",

  "East Sands",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f662a74fb848dae79f6a46_WhatsApp%20Image%202025-10-20%20at%2017.24.27_a9ec7b97.jpg",

  "Kilve Beach",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7ae6ad6b9191fe5bbf189_20250905_142406%20-%20Athena%20Allen%20(1).jpg",

  "Beadnell Haven",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f663d2083ef94524503f95_team%20compressed.jpg",

  "Lee-on-the-Solent",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f7a41bfa2bbfcef8002094_IMG_20251018_142731.jpg",

  "Shoeburyness",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f664397fe026047360f36a_shoeburyness%20team.jpg"
)
```

::: {style="height:20px;"}
:::

::: {style="height:20px;"}
:::

During BioBlitz Battle games, all species recorded earn a score. These scores reflect how often this species is recorded in the overall Big Rock Pool Challenge dataset. The higher scoring species are those that are recorded less often within the project and may represent more unusual or less commonly observed species. 

In this report, scores are given between 1-100, according to our current scoring system 2.1


The highest scoring species recorded during this period are presented below.

```{r top-species-all-data}
all_hub_top_species <- all_iconic %>%
  group_by(hub, scientific_name) %>%
  summarise(
    Score = max(Score, na.rm = TRUE),
    photo = first(na.omit(taxon.default_photo.medium_url)),
    photo_attribution = first(na.omit(taxon.default_photo.attribution)),
    photo_license = first(na.omit(taxon.default_photo.license_code)),
    recorder = first(user.login),
    .groups = "drop"
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok,
      photo_attribution,
      ""
    )
  ) %>%

  arrange(hub, desc(Score)) %>%
  group_by(hub) %>%
  slice_head(n = 3) %>%

  mutate(
    img_html = paste0(

      "<div style='
        flex:1;
        min-width:180px;
        max-width:220px;
        background:#ffffff;
        border-radius:14px;
        padding:14px;
        box-shadow:0 2px 8px rgba(0,0,0,0.08);
        border:1px solid #ececec;
        text-align:center;
      '>",

      "<img src='", photo,
      "' style='
          width:100%;
          height:140px;
          object-fit:cover;
          border-radius:10px;
          margin-bottom:10px;
      '>",

      "<div style='font-size:10px;color:#777;margin-bottom:8px;'>",
      attribution,
      "</div>",

      "<div style='font-style:italic;font-size:14px;margin-bottom:6px;'>",
      scientific_name,
      "</div>",

      "<div style='font-size:12px;color:#666;margin-bottom:6px;'>",
      "Rarity score: ", round(Score, 1),
      "</div>",

      "<div style='font-size:11px;color:#444;'>",
      "<strong>Observed by</strong><br>",
      recorder,
      "</div>",

      "</div>"
    )
  ) %>%

  summarise(
    top_species_html = paste0(
      "<div style='display:flex;gap:16px;justify-content:center;flex-wrap:wrap;margin-top:12px;'>",
      paste(img_html, collapse = ""),
      "</div>"
    ),
    .groups = "drop"
  )
```

```{r notable_species}

notable_species <- q1_iconic %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    Score = max(Score, na.rm = TRUE),
    observations = n(),
    photo = first(na.omit(taxon.default_photo.medium_url)),
    recorder = first(user.login),
    hub = first(hub),
    .groups = "drop"
  ) %>%
  arrange(desc(Score)) %>%
  slice_head(n = 10)
```

```{r hub-iconic-plots, echo=FALSE, warning=FALSE, message=FALSE}

dir.create("hub_plots", showWarnings = FALSE)

hub_iconic_small <- q1_iconic %>%
  count(hub, iconic_group) %>%
  group_by(hub) %>%
  mutate(prop = n / sum(n)) %>%
  left_join(icon_lookup, by = "iconic_group")

for(h in unique(hub_iconic_small$hub)) {

  dat <- hub_iconic_small %>%
    filter(hub == h)

  p <- ggplot(
    dat,
    aes(
  x = reorder(iconic_group, prop),
  y = prop
    )
  ) +

   geom_col(
  fill = "#9ecae1",
  show.legend = FALSE
) +

    geom_image(
      aes(image = icon_path),
      by = "width",
      size = 0.05,
      nudge_y = max(dat$prop) * 0.03
    ) +

    expand_limits(
      y = max(dat$prop) * 1.12
    ) +

    coord_flip() +

    scale_y_continuous(labels = scales::percent) +

    labs(
      x = NULL,
      y = NULL
    ) +

    theme_minimal(base_size = 10) +

    theme(
      plot.margin = margin(2,2,2,2),
      panel.grid.minor = element_blank(),
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(size = 7)
    )

  ggsave(
    filename = paste0("hub_plots/", safe_id(h), ".jpg"),
    plot = p,
    width = 4,
    height = 2.8,
    dpi = 120
  )
}
```

```{r all-notable-species, results='asis', echo=FALSE}
# ---------------------------------------
# All-time standout species
 all_time_top <- BRPC_dat %>%
  left_join(
    rarity_scores,
    by = c("scientific_name" = "name")
  ) %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    Score = max(Score, na.rm = TRUE),

    photo = first(na.omit(taxon.default_photo.medium_url)),
    photo_attribution = first(na.omit(taxon.default_photo.attribution)),
    photo_license = first(na.omit(taxon.default_photo.license_code)),

    recorder = first(user.login),
    hub = first(hub),

    .groups = "drop"
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok,
      photo_attribution,
      ""
    )
  ) %>%

  arrange(desc(Score)) %>%
  slice_head(n = 3)

```

```{r standout-species-render, results='asis', echo=FALSE}
library(htmltools)

render_species_cards <- function(dat, title_text) {

  cards <- lapply(seq_len(nrow(dat)), function(i) {

    row <- dat[i, ]

    div(
      style = "
        flex:1;
        min-width:260px;
        max-width:320px;
        background:#ffffff;
        border-radius:18px;
        padding:20px;
        box-shadow:0 3px 12px rgba(0,0,0,0.08);
        border:1px solid #e6e6e6;
        text-align:center;
      ",

tags$a(
  href = row$photo,
  target = "_blank",
  tags$div(
    style = sprintf(
      "width:100%%;height:140px;border-radius:10px;background-image:url('%s');background-size:cover;background-position:center;",
      row$photo
    )
  )
),

      # FIX: must come from row, not global
      tags$div(
        style = "font-size:10px;color:#777;margin-top:6px;",
        HTML(row$attribution %||% "")
      ),

      div(
        style = "
          display:inline-block;
          padding:8px 14px;
          border-radius:999px;
          background:#f5f5f5;
          font-size:15px;
          margin-bottom:14px;
        ",
        strong("Score: "),
        round(row$Score, 1)
      ),

      tags$p(
        style = "margin-bottom:6px;",
        strong("Observed by"), tags$br(),
        row$recorder
      ),

      tags$p(
        style = "margin-top:0;",
        strong("Hub"), tags$br(),
        row$hub
      )
    )
  })

  tagList(
    tags$h2(
      style = "margin-top:40px;margin-bottom:24px;text-align:center;",
      title_text
    ),
    div(
      style = "
        display:flex;
        gap:24px;
        justify-content:center;
        flex-wrap:wrap;
        margin-bottom:50px;
      ",
      cards
    )
  )
}

render_species_cards(
  all_time_top,
  "Highest Scoring Species Across All Project Data"
)
```

```{r all-hub-report-join}
all_hub_report <- all_hub_summaries %>%
  left_join(all_hub_top_recorders, by = "hub") %>%
  left_join(all_hub_top_identifiers, by = "hub") %>%
  left_join(all_iconic %>%
              group_by(hub) %>%
              summarise(top_groups = paste(names(sort(table(iconic_group), decreasing = TRUE)[1:3]), collapse = ", "),
                        .groups = "drop"),
            by = "hub") %>%
  left_join(all_hub_top_species, by = "hub") %>%
  left_join(hub_logos, by = "hub") %>%
  left_join(hub_site_photos, by = "hub") %>%
  distinct(hub, .keep_all = TRUE)
```

::: {style="height:20px;"}
:::

### Hub Summaries

A breakdown of records collected at each Big Rock Pool Challenge hub is presented below, including key contributors, iconic taxa grouping, and top scoring species.

```{r hubs-render, results='asis', echo=FALSE}
library(htmltools)
library(dplyr)
library(scales)

default_avatar <- "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/68c02913eb9eedda1582018a_iNaturalist%20app.png"

hubs <- all_hub_report$hub
hubs <- hubs[!is.na(hubs)]

out <- lapply(hubs, function(h) {

  row <- all_hub_report %>%
    filter(hub == h) %>%
    slice(1)

  div(

    style = "
      background:#ffffff;
      border-radius:20px;
      padding:28px;
      margin-bottom:36px;
      box-shadow:0 4px 14px rgba(0,0,0,0.08);
      border:1px solid #e8e8e8;
      width:100%;
      box-sizing:border-box;
      overflow:hidden;
    ",

    # ================= HEADER =================
    div(
      style = "text-align:center;margin-bottom:24px;",

      img(
        src = row$logo_url,
        onerror = paste0("this.src='", default_avatar, "';"),
        style = "
          width:180px;
          height:180px;
          object-fit:contain;
          margin-bottom:14px;
        "
      ),

      h2(h, style = "margin:0;font-size:28px;"),

      div(
        style = "
          display:inline-block;
          margin-top:14px;
          padding:10px 18px;
          border-radius:999px;
          background:#f5f5f5;
          font-size:14px;
        ",
        strong(comma(row$records)), " observations · ",
        strong(comma(row$contributors)), " contributors · ",
        strong(comma(row$species)), " species"
      )
    ),

    # ================= GRID =================
    div(
      style = "
        display:grid;
        grid-template-columns:1fr 1fr;
        gap:22px;
        width:100%;
      ",

      # ---- SITE PHOTO ----
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",

        h3( style = "margin-top:0;text-align:center;"),

        img(
          src = row$site_photo,
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            height:320px;
            object-fit:cover;
            border-radius:12px;
          "
        )
      ),

      # ---- ICONIC TAXA ----
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
          text-align:center;
        ",

        h3("Iconic Taxa Composition", style = "margin-top:0;"),

        img(
          src = paste0("hub_plots/", safe_id(h), ".jpg"),
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            max-width:480px;
            border-radius:10px;
          "
        ),

        tags$p(
          style = "margin-top:16px;font-size:13px;color:#555;",
          strong("Top iconic groups: "),
          row$top_groups
        )
      ),

      # ---- TOP RECORDERS ----
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",

        h3("Top Recorders", style = "text-align:center;"),

        div(
          style = "
            display:flex;
            gap:16px;
            justify-content:center;
            flex-wrap:wrap;
          ",
          HTML(row$top_recorders_html)
        )
      ),

      # ---- TOP IDENTIFIERS ----
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",

        h3("Top Identifiers", style = "text-align:center;"),

        div(
          style = "
            display:flex;
            gap:16px;
            justify-content:center;
            flex-wrap:wrap;
          ",
          HTML(row$top_identifiers_html)
        )
      )
    ),

    # ================= SPECIES =================
    div(
      style = "margin-top:28px;",

      h3("Top Scoring Species", style = "text-align:center;margin-bottom:18px;"),

      HTML(row$top_species_html)
    )
  )
})

tagList(out)
```

## Quarter 1 2026 Data {#q1-2026}

This section presents only records from Quarter 1 of 2026 (January–March) to examine early-year biodiversity patterns across monitoring hubs. Seasonal variation is expected in intertidal ecosystems due to changes in temperature, storm exposure, and species life cycles.

```{r q1-summary}
tagList(

  render_summary_card(
    "Quarter 1 2026 Summary",
    observations = nrow(q1_base),
    contributors = n_distinct(q1_base$user.login),
    species = n_distinct(na.omit(q1_base$scientific_name))
  )
)

```

```{r iconic-q1-plot}
ggplot(
  iconic_summary,
  aes(
    x = reorder(iconic_group, n),
    y = n,
    fill = iconic_group
  )
) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Iconic Taxa Recorded During Q1 2026",
    x = NULL,
    y = "Observations"
  )
```

```{r q1-hub-recorder-summary}

q1_recorders <- q1_base %>%
  filter(!is.na(user.login)) %>%
  group_by(user.login) %>%
  summarise(
    records = n(),
    icon = first(na.omit(user.icon_url)),
    .groups = "drop"
  ) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    records = comma(records),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          margin-bottom:10px;
          border:2px solid #ececec;
          object-fit:cover;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      records,
      " records</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Q1 Top Recorders", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(q1_recorders$card_html, collapse = ""))
  )
)
```

```{r q1-top-identifiers}

q1_identifiers <- q1_base %>%
  filter(!is.na(identifications)) %>%
  select(identifications) %>%
  mutate(
    identifier = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.login" %in% names(.x)) {
        character(0)
      } else {
        .x$user.login
      }
    }),
    icon = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.icon_url" %in% names(.x)) {
        character(0)
      } else {
        .x$user.icon_url
      }
    })
  ) %>%
  select(identifier, icon) %>%
  unnest(c(identifier, icon)) %>%
  filter(!is.na(identifier), identifier != "") %>%
  group_by(identifier) %>%
  summarise(
    identifications = n(),
    icon = first(na.omit(icon)),
    .groups = "drop"
  ) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    identifications = comma(identifications),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          margin-bottom:10px;
          border:2px solid #ececec;
          object-fit:cover;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      identifier,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      identifications,
      " identifications</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Q1 Top Identifiers", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(q1_identifiers$card_html, collapse = ""))
  )
)
```

```{r q1-hub-summary}

q1_hub_summaries <- q1_iconic %>%
  group_by(hub) %>%
  summarise(
    records = n(),
    contributors = n_distinct(user.login),
    species = n_distinct(scientific_name),
    .groups = "drop"
  )

```

```{r q1-hub-top-groups}

q1_hub_top_groups <- q1_iconic %>%
  count(hub, iconic_group, sort = TRUE) %>%
  group_by(hub) %>%
  slice_head(n = 3) %>%
  summarise(
    top_groups = paste(iconic_group, collapse = ", "),
    .groups = "drop"
  )

```

```{r q1-hub-top-species}
q1_hub_top_species <- q1_iconic %>%
  filter(!is.na(Score)) %>%

  group_by(hub, scientific_name) %>%

  arrange(desc(Score)) %>%

  slice(1) %>%

  ungroup() %>%

  transmute(
    hub,
    scientific_name,
    Score,

    photo = taxon.default_photo.medium_url,
    photo_attribution = taxon.default_photo.attribution,
    photo_license = taxon.default_photo.license_code,

    recorder = user.login
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok &
      !is.na(photo_attribution) &
      photo_attribution != "",
      photo_attribution,
      ""
    )
  ) %>%

  arrange(hub, desc(Score)) %>%

  group_by(hub) %>%

  slice_head(n = 3) %>%

  mutate(
    img_html = paste0(
      "<div style='
        flex:1;
        min-width:180px;
        max-width:220px;
        background:#ffffff;
        border-radius:14px;
        padding:14px;
        box-shadow:0 2px 8px rgba(0,0,0,0.08);
        border:1px solid #ececec;
        text-align:center;
      '>",

      "<img src='", photo, "' style='
          width:100%;
          height:140px;
          object-fit:cover;
          border-radius:10px;
          margin-bottom:6px;
      '>",

      "<div style='font-size:10px;color:#777;margin-bottom:8px;'>",
      attribution,
      "</div>",

      "<div style='font-style:italic;font-size:14px;margin-bottom:8px;'>",
      scientific_name,
      "</div>",

      "<div style='font-size:12px;color:#666;margin-bottom:8px;'>",
      "Rarity score: ", round(Score, 1),
      "</div>",

      "<div style='font-size:12px;color:#444;line-height:1.4;'>",
      "<strong>Observed by</strong><br>",
      recorder,
      "</div>",

      "</div>"
    )
  ) %>%

  summarise(
    top_species_html = paste0(
      "<div style='display:flex;gap:16px;justify-content:center;flex-wrap:wrap;margin-top:12px;'>",
      paste(img_html, collapse = ""),
      "</div>"
    ),
    .groups = "drop"
  )
```

```{r q1-hub-top-recorders}

q1_hub_top_recorders <- q1_iconic %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    records = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(

    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(records),
      " observations</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_recorders_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )

```

```{r q1-hub-top-identifiers}

q1_hub_top_identifiers <- q1_iconic %>%
  filter(!is.na(identifications)) %>%
  select(hub, identifications) %>%
  tidyr::unnest(identifications) %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    identifications = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(

    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(identifications),
      " identifications</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_identifiers_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )

```

```{r q1-top-species-card}
q1_top <- q1_iconic %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    n = n(),

    Score = max(Score, na.rm = TRUE),

    photo = first(na.omit(taxon.default_photo.medium_url)),
    photo_attribution = first(na.omit(taxon.default_photo.attribution)),
    photo_license = first(na.omit(taxon.default_photo.license_code)),

    recorder = first(user.login),
    hub = first(hub),

    .groups = "drop"
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok,
      photo_attribution,
      ""
    )
  ) %>%

  arrange(desc(Score)) %>%
  slice_head(n = 3)

```

```{r q1-species-render, results='asis', echo=FALSE}
render_species_cards(
  q1_top,
  "Highest Scoring Species Recorded During Q1 2026"
)
```

```{r q1-hub-report}

q1_hub_report <- q1_hub_summaries %>%
  left_join(q1_hub_top_recorders, by = "hub") %>%
  left_join(q1_hub_top_identifiers, by = "hub") %>%
  left_join(q1_hub_top_groups, by = "hub") %>%
  left_join(q1_hub_top_species, by = "hub") %>%
  left_join(hub_logos, by = "hub") %>%
  left_join(hub_site_photos, by = "hub") %>%
  distinct(hub, .keep_all = TRUE)

```

### Hub Summaries Q1

```{r q1-hub-render}
q1_hubs <- q1_hub_report$hub
q1_hubs <- q1_hubs[!is.na(q1_hubs)]

out <- lapply(q1_hubs, function(h) {

  row <- q1_hub_report %>%
    filter(hub == h) %>%
    slice(1)

  div(
    style = "
      background:#ffffff;
      border-radius:20px;
      padding:28px;
      margin-bottom:36px;
      box-shadow:0 4px 14px rgba(0,0,0,0.08);
      border:1px solid #e8e8e8;
      width:100%;
      box-sizing:border-box;
      overflow:hidden;
    "
    ,

    # HEADER
    div(
      style = "text-align:center;margin-bottom:24px;",

      img(
        src = row$logo_url,
        onerror = paste0("this.src='", default_avatar, "';"),
        style = "
          width:180px;
          height:180px;
          object-fit:contain;
          margin-bottom:14px;
        "
      ),

      h2(h, style = "margin:0;font-size:28px;"),

tags$div(
  style = "
    font-size:13px;
    color:#666;
    margin-top:6px;
    margin-bottom:4px;
  ",
  "Quarter 1: 01/01/2026–31/03/2026"
),

      div(
        style = "
          display:inline-block;
          margin-top:14px;
          padding:10px 18px;
          border-radius:999px;
          background:#f5f5f5;
          font-size:14px;
        ",
        strong(comma(row$records)), " observations · ",
        strong(comma(row$contributors)), " contributors · ",
        strong(comma(row$species)), " species"
      )
    ),

    # GRID
    div(
      style = "
        display:grid;
        grid-template-columns:1fr 1fr;
        gap:22px;
        width:100%;
      ",

      # SITE PHOTO
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",

        img(
          src = row$site_photo,
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            height:320px;
            object-fit:cover;
            border-radius:12px;
          "
        )
        ),

      # ICONIC TAXA
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
          text-align:center;
        ",

        h3("Iconic Taxa Composition", style = "margin-top:0;"),

        img(
          src = paste0("hub_plots/", safe_id(h), ".jpg"),
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            max-width:480px;
            border-radius:10px;
          "
        ),

        tags$p(
          style = "margin-top:16px;font-size:13px;color:#555;",
          strong("Top iconic groups: "),
          row$top_groups
        )
      ),

      # TOP RECORDERS
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",
        h3("Top Recorders", style = "text-align:center;"),
        div(style="display:flex;gap:16px;justify-content:center;flex-wrap:wrap;",
            HTML(row$top_recorders_html))
      ),

      # TOP IDENTIFIERS
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",
        h3("Top Identifiers", style = "text-align:center;"),
        div(style="display:flex;gap:16px;justify-content:center;flex-wrap:wrap;",
            HTML(row$top_identifiers_html))
      )
    ),

    # SPECIES
    div(
      style = "margin-top:28px;",
      h3("Top Scoring Species", style = "text-align:center;margin-bottom:18px;"),
      HTML(row$top_species_html)
    )
  )
})

tagList(out)
```
